\thetitle

### 2021-02-28 writing/challenges.tex
organize sections/titles/labels in paper
factor mutate
revise intro to point to new 'why rational' section

\section{Why Is It Hard to Evaluate Blame?}
\label{sec:challenges}

@@

The general challenge of evaluating blame is a methodological one. Unlike most
current research on programming languages, the question seems to call for
empirical research similar to that of human-computer interaction. At the same
time, a significant evaluation result demands a large amount of
data. As~\citet{lksfd-popl-2020} recently demonstrated, the way around this
dilemma is to postulate the {\em rational programmer\/}---the equivalent of a
{\it homo economicus\/} for programming languages---and to implement it as an algorithm.

In the context of gradual typing, the rational programmer translates the
Wadler-Findler slogan into a debugging method, searching for the source of the
impedance mismatch in an incremental fashion. Measuring this simulated
programmer's behavior on a large number of debugging scenarios yields data that
is similar to data collected in an empirical, human-facing manner.

@@

Thus, an experimental analysis of blame needs a mostly new set of mutators.
Roughly speaking, the new mutators inject type errors into fully typed programs.
Applying such a mutator to any typed component produces a mutatnt component.  A
debugging scenario results from removing the types from the mutatnt. For the
design of such mutators, the authors relied on their own extensive programming
experience though not without discovering a major pitfall: some of their
original mutators systematically produced programs that immediately revealed the
source of the impedance mismatch. About two thirds of the original mutators
survived this first scrutiny testing. All of the remaining ones yield {\em
interesting debugging scenarios\/} (see section~\ref{sub:interesting}).

>>>>

@@

@@

Thus, an experimental analysis of blame needs a mostly new set of mutators.
Roughly speaking, the new mutators inject type errors into fully typed programs.
Applying such a mutator to any typed component produces a mutatnt component.  A
debugging scenario results from removing the types from the mutatnt. For the
design of such mutators, the authors relied on their own extensive programming
experience though not without discovering a major pitfall: some of their
original mutators systematically produced programs that immediately revealed the
source of the impedance mismatch. About two thirds of the original mutators
survived this first scrutiny testing. All of the remaining ones yield {\em
interesting debugging scenarios\/} (see section~\ref{sub:mutate-interesting}).

### 2021-02-28 writing/introduction.tex
organize sections/titles/labels in paper
revise intro to point to new 'why rational' section

\section{Does Blame Matter?}
\label{sec:introduction}

@@

This contrast between theoretical research and industrial applications of gradual typing raises the question 
\begin{quote}
 \it
 whether blame assignment adds any value to a gradually typed language,
 especially for the benefit of the working programmer.
\end{quote}
Given the long-standing academic interest in blame and its complete absence in
industrial systems, it comes as an even bigger surprise that the research
literature and the industrial blog world do not discuss any possible answers.
Instead, when language designers make relevant decisions, they seem to answer it
one way or another without any scientific justification. Then again, the
community has thus far failed to offer a method for evaluating blame assignment.

This paper's first contribution is {\em an automated method for evaluating the
effectiveness of blame assignment strategies\/} in the gradual typing world.
speaking, the method simulates a programmer who follows the Wadler-Findler
slogan quoted above; its automation borrows elements of of Lazarek et
al.~\cite{lksfd-popl-2020}'s recent work.\footnote{\citet{cc-snapl-19} use this
tracing method to localize type errors reported by gradual typing systems based
on inference.}  Such a \emph{rational programmer} can find impedance-mismatch
bugs by adding types to a partially typed program (see
section~\ref{sec:landscape}). That is, when the run-time system signals an
impedance mismatch, a good blame message helps the programmer infer where to add
an additional type specification because, by the basic theory, well-typed
components cannot cause impedance mismatches. Otherwise, if the additional type
does {\em not\/} cause a static type error, the program is run again to signal
another violation. Then the programmer can heed blame one more time to add yet
another type specification. Evaluating the quality of blame assignment thus
becomes a question of whether the tracing the error messages finds the source of
the mistake and how many steps it takes to get to this point.

@@

The results of the method's first application (see section~\ref{sec:discussion})
are at least somewhat surprising. In principle they validate the conjectures
behind the work of theoreticians. First, a good blame assignment strategy helps
with the search for impedance mismatches between the specified types and the
behavior of untyped components.  Second, Natural's wrapper-based blame tracking
is more useful than Transient's ``collective blame'' tracking algorithm, which
in turn is superior to Erasure. {\em But\/}, the application of the method also
indicates problems with the expectations of theoreticians. The first problem is
that the cost of Transient's blame can be huge. The second one concerns the
difference between the methods; neither is {\it Natural\/} vastly better than
{\it Transient\/} nor are the two blame assignment methods clearly preferable to
{\it Erasure\/}. The third one casts doubt on the current the trade-off between
the cost of blame-assignment systems and the quality of their blame messages.
In turn, these problems suggests that, on one hand, the existing theory does not
predict practice properly, and on the other hand, the existing practice needs to
be supported by additional data.

>>>>

@@

This contrast between theory and applications of gradual typing raises the question 
\begin{quote}
 \it
 whether blame assignment adds any value to a gradually typed language,
 especially for the benefit of the working programmer.
\end{quote}
Given the long-standing academic interest in blame and its complete absence in
industrial systems, it comes as an even bigger surprise that the research
literature and the industrial blog world do not discuss any possible answers.
Instead, when language designers make relevant decisions, they seem to answer it
one way or another without any scientific justification. Then again, the
community has thus far failed to offer a method for evaluating blame assignment.

This paper's first contribution is {\em an automated method for evaluating the
effectiveness of blame assignment strategies\/} in the gradual typing world.
The top-level innovation is the idea of a {\em rational programmer\/}, that is,
a programmer that acts only in response to available information (see section~\ref{sec:why-rational}). In the case
of an impedance mismatch, the available information consists of the error
message and the current state of the program. The rational programmer can hence
use the former to change the latter---and this systematic, information-driven
process can be implemented.\footnote{\citet{cc-snapl-19} use this tracing
method to localize type errors reported by gradual typing systems based on
inference.} Implementing it means overcoming three major challenges: injecting
representative impedance mismatches; putting the various kinds of error
information to comparable use; and sampling the huge space of possibilities
properly (see section~\ref{sec:challenges}).

% The method simulates a programmer who follows the Wadler-Findler slogan quoted
% above; its automation borrows elements of of Lazarek et
% al.~\cite{lksfd-popl-2020}'s recent work.  Such a \emph{rational programmer} can find impedance-mismatch
% bugs by adding types to a partially typed program (see
% section~\ref{sec:landscape}). That is, when the run-time system signals an
% impedance mismatch, a good blame message helps the programmer infer where to
% add an additional type specification because, by the basic theory, well-typed
% components cannot cause impedance mismatches. Otherwise, if the additional type
% does {\em not\/} cause a static type error, the program is run again to signal
% another violation. Then the programmer can heed blame one more time to add yet
% another type specification. Evaluating the quality of blame assignment thus
% becomes a question of whether the tracing the error messages finds the source
% of the mistake and how many steps it takes to get to this point.

@@

The results of the method's first application (see section~\ref{sec:discussion})
are at least somewhat surprising. In principle they validate the conjectures
behind the work of theoreticians. First, a good blame assignment strategy helps
with the search for impedance mismatches between the specified types and the
behavior of untyped components.  Second, Natural's wrapper-based blame tracking
is more useful than Transient's ``collective blame'' tracking algorithm, which
in turn is superior to Erasure. {\em But\/}, the application of the method also
indicates problems with the expectations of theoreticians. The first problem is
that the cost of Transient's blame can be huge. The second one concerns the
difference between the methods; neither is {\it Natural\/} vastly better than
{\it Transient\/} nor are the two blame assignment methods clearly preferable to
{\it Erasure\/}.  In turn, these problems suggests that, on one hand, the
existing theory does not predict practice properly, and on the other hand, the
existing practice may need to find additional ways to collect data.
